Influence of Time Integration Methodology on Double Detonation Type Ia Supernova Nucleosynthesis

### 2023-06-14 subch_tols/paper.tex
add a bit more

@@

The reaction system is integrating using a version of the VODE ODE integrator
\citep{vode} ported to C++.  Our modifications to this integrator are described
in \citet{castro_simple_sdc}.

We will run a suite of simulations all using the same initial model.
The initial model is exactly the same as in \citep{castro_simple_sdc}.

@@

@@

In the near future we will show how to include various nuclear
statistical equilibrium approximations into the simplified-SDC
formalism while retaining second-order accuracy.

>>>>

All simulations use the same network, a 22 nucleus network designed for
He and C burning, using the ReacLib rates \cite{reaclib} and produced with
\pynucastro \citep{pynucastro2}.  This network, called {\tt subch\_simple} was
described in \cite{zhi2023}.

The reaction system is integrated using a version of the VODE ODE integrator
\citep{vode} ported to C++.  Our modifications to this integrator are described
in \citet{castro_simple_sdc}.

@@

We will run a suite of simulations all using the same initial model.
The initial model is constructed following the methodology in
\citet{subchandra}.  We use an isothermal core of C/O of $1.1~M_\odot$
with $T = 10^7~\mathrm{K}$ and a thin transition region where the
temperature ramps up to $1.75\times 10^8~\mathrm{K}$ and the
composition changes to 99\% \isot{He}{4} and 1\% \isot{N}{14}.  This
added \isot{N}{14} helps produce protons via
$\isotm{N}{14}(\alpha,\gamma)\isotm{F}{18}(\alpha,p)\isotm{Ne}{21}$
which then allow for
$\isotm{C}{12}(p,\gamma)\isotm{N}{13}(\alpha,p)\isotm{O}{16}$.  This
sequence can be faster than
$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$, as pointed out by
\citet{shenbildsten} and is important for getting the detonation speed
correctly.  This He envelope is then integrated isentropically.  The
density of the transition from the underlying CO white dwarf to the He
envelope was selected to yield an envelope mass of $0.05~M_\odot$.

We place a small temperature perturbation in the He layer using the
same form as in \citet{castro_simple_sdc}:
\begin{equation}
  T = T_0 \left [ 1 + X(\isotm{He}{4}) f (1 + \tanh(2 - r_1)) \right ]
\end{equation}
where
\begin{equation}
  r_1 = \left [ x^2 + (y - r_0)^2 \right ]^{1/2} / \lambda
\end{equation}
and
\begin{equation}
  r_0 = r_\mathrm{pert} + r_\mathrm{base}
\end{equation}
Here, $r_\mathrm{base}$ is the radius at which the helium layer begins
and $r_\mathrm{pert}$ is the distance above the base to put the
perturbation.  We choose $r_\mathrm{pert} = 100~\mathrm{km}$.  The
temperature is perturbed above the initial model value, denoted as
$T_0$ here.  The amplitude of the perturbation is $f = 3$ and the
scale of the perturbation is $\lambda = 12.5~\mathrm{km}$, and
$r_\mathrm{pert} = 100~\mathrm{km}$.  This is a very small
perturbation, but enough to seed the initial He detonation in the
envelope.

@@

\begin{figure*}[t]
\centering
\plotone{subch_Temp_sequence}
\caption{\label{fig:temp_sequence} Time-sequence of the SDC run showing the temperature.}
\end{figure*}

\begin{figure*}[t]
\centering
\plotone{subch_lap_rho_sequence}
\caption{\label{fig:lap_rho_sequence} Time-sequence of the SDC run showing the compression.}
\end{figure*}

@@

The ignition of the carbon detonation remains numerically challenging for the ODE integrator, especially with large timesteps.  \MarginPar{add some info on why it fails}
In the near future we will show how to include various nuclear
statistical equilibrium approximations into the simplified-SDC
formalism while retaining second-order accuracy.
